---
title: "simulation"
author: "Mohamed"
date: "24/12/2020"
output:
  pdf_document: default
  html_document: default
---

## Introduction 


The purpose of this data analysis is to investigate the exponential distribution
and compare it to the Central Limit Theorem. For this analysis, the lambda will 
be set to 0.2 for all of the simulations. This investigation will compare the 
distribution of averages of 40 exponentials over 1000 simulations.


## Load library

```{r ,ECHO=TRUE}
library(tinytex)
library(ggplot2)
```
## Simulations
Set the simulation variables lambda, exponentials, and seed.
```{r ,ECHO=TRUE}


library(ggplot2)

set.seed(28) ##to ensure reproducibility, I am setting seed arbitrary on 28.

lambda <- 0.2

```

Run Simulations with variables
```{r ,ECHO=TRUE}
exp_means =NULL
for (i in 1 : 1000) 
  exp_means = c(exp_means , mean(rexp(40, 0.2)))
```




## Drowing histogram to show simple mean
#### Sample Mean
Calculating the mean from the simulations with give the sample mean.
```{r}
hist(exp_means)

simMeanMean <- mean(exp_means)
# theoretical exponential mean
theMean <- 1/lambda

abline(v=theMean, col="red", lwd=3)

```



## Gaussian distribution
Gaussian distribution  compare with a normal distribution.
```{r}
hist(exp_means, breaks=20, prob=TRUE, xlab="Mean of exponentials", ylab="Frequency", 
     col="IndianRed3")
curve(dnorm(x, mean=mean(exp_means), sd=sd(exp_means)), col="blue", lwd=2, 
      lty = "dotted", add=TRUE, yaxt="n")
curve(dnorm(x, mean=5, sd=0.79), col="black", lwd=2, add=TRUE, yaxt="n")
```
